### 2014-01-20 paper/manuscript.tex
Second version of the paper, with figures.

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{times}
\usepackage{graphicx}
\usepackage{color}
\usepackage{multirow}
%
\usepackage[authoryear]{natbib}
%
\usepackage{rotating}
\usepackage{bbm}
\usepackage{latexsym}
%\DeclareGraphicsExtensions{.eps,.png}

%%% margins 
\textheight 23.4cm
\textwidth 14.65cm
\oddsidemargin 0.375in
\evensidemargin 0.375in
\topmargin  -0.55in
%
\renewcommand{\baselinestretch}{2}
%
\interfootnotelinepenalty=10000
%
\renewcommand{\thesubsubsection}{\arabic{section}.\arabic{subsubsection}}
\newcommand{\myparagraph}[1]{\ \\{\em #1}.\ \ }
\newcommand{\citealtt}[1]{\citeauthor{#1},\citeyear{#1}}
\newcommand{\myycite}[1]{\citep{#1}}

@@

\newcommand{\ie}{\emph{i.e., }}
\newcommand{\eg}{\emph{e.g., }}
%% \newcommand{\cf}{\emph{cf. }}
%% \newcommand{\etc}{\emph{etc.}}
%% \newcommand{\wrt}{w.r.t. }

\usepackage{nameref}
\newcommand{\fig}[1]{Figure \ref{fig:#1}}
\newcommand{\figlabel}[1]{\label{fig:#1}}
\newcommand{\eqn}[1]{Equation \ref{eqn:#1}}
\newcommand{\eqnlabel}[1]{\label{eqn:#1}}
\newcommand{\scn}[1]{\textsc{\nameref{scn:#1}}}
\newcommand{\scnlabel}[1]{\label{scn:#1}}

\renewcommand{\thefootnote}{\normalsize \arabic{footnote}}

\begin{document}
\hspace{13.9cm}1

@@

{\LARGE Heterogeneity increases information transmission of neuronal populations}

\ \\
{\bf \large Eric G. Hunsberger$^{\displaystyle 1}$, Matthew Scott$^{\displaystyle 2}$, Chris Eliasmith$^{\displaystyle 1}$}\\
{$^{\displaystyle 1}$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Ontario, Canada.}\\
{$^{\displaystyle 2}$Department of Applied Mathematics, University of Waterloo, Waterloo, Ontario, Canada.}\\
%

@@

\thispagestyle{empty}
\markboth{}{NC instructions}
%
\ \vspace{-0mm}\\
%
%Abstract
\begin{center} {\bf Abstract} \end{center}
Noise---specifically random fluctuations added to the membrane voltages of a neuronal population---has been shown to increase the amount of information a neuronal population can encode about an input signal; this phenomenon is one type of stochastic resonance. We demonstrate that heterogeneity within a population of simulated neurons---specifically the random variation of membrane bias currents across a population---produces similar benefits as additive noise in terms of increasing information encoding within the population. Furthermore, we find that both noise and heterogeneity use a common set of three mechanisms to increase information transmission: 1) they temporally desynchronize the firing of neurons within the population, 2) they decrease the response time of the population of neurons to sudden changes in the input signal, and 3) they linearize the response of a population to a stimulus. We examine each mechanism in detail, describing how these mechanisms operate in neuronal populations, and demonstrating how they operate in populations of simulated FitzHugh-Nagumo and leaky integrate-and-fire neurons. These three mechanisms work to distribute information equally across all neurons in the population, both in terms of signal timing (desynchronization and response time) and in terms of signal amplitude (linearization). Without noise or heterogeneity, all neurons in the population encode the same aspects of the input signal; adding noise or heterogeneity allows all neurons to encode \emph{complementary} aspects of the input signal, thereby preserving more information. Overall, the simulations detailed in this paper highlight the importance of heterogeneity in neuronal population coding, and demonstrate that it can be equally effective as stochastic resonance in increasing information transmission. 
%%%%%%%%%%%

\section{Introduction}
\scnlabel{introduction}

@@

The early stochastic resonance literature dealt mainly with \emph{subthreshold} stochastic resonance \citep{Gammaitoni1998}. In a deterministic setting, systems which only respond to inputs above a certain threshold---such as neurons---cannot transmit any information about subthreshold input signals, \ie input signals which are entirely below the system's threshold. In contrast, a noisy input signal (\ie a signal with an added stochastic offset) will occasionally cross the system's threshold, and thereby allows the system to transmit more information about the input signal. Too much noise, though, causes the input signal to become lost in random fluctuations. A moderate amount of noise---not too much, and not too little---is needed to optimize information transmission by the system about the input signal \citep{Wiesenfeld1994,Longtin1998}.

A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009}. In neuroscience, suprathreshold SR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s). This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units, and found that noise can increase information transmission in such systems. Subsequent research has examined the SSR effect in populations of more sophisticated simulated neurons \citep{Stocks2001}.

Heterogeneity among neurons---the fact that not all neurons are identical, and vary across a wide range of parameters---is a topic that has largely been ignored in the stochastic resonance literature, and neuroscience literature in general \citep{Machens2010}. It is well known that there are large differences between neurons in different parts of the nervous system, and that there are many different types of neurons, but research into the role of heterogeneity within a group of neurons of the same type is limited (\cite{Eliasmith2003} is a notable exception). In the field of stochastic resonance, the effects of noise in a heterogeneous population have been examined, but generally in specific cases such as for input signals much larger than the range of heterogeneity \citep{Stocks2000}, or for very large levels of noise \citep{McDonnell2006}. Furthermore, research into stochastic resonance with heterogeneity has focused on populations of binary threshold units, which fail to capture the dynamics of actual neurons.

The purpose of this paper is to compare the benefits of noise and the benefits of heterogeneity on information transmission by a population of simulated neurons. While previous studies have looked at the effects of noise, often claiming that noise is the central mechanism of SR and SSR effects \citep{Collins1995,Stocks2001} (see \cite{McDonnell2011} for a review), we are not aware of any work that has examined the effects of heterogeneity in similar circumstances. As we show, many effects of noise can also be achieved with heterogeneity. We argue that this is at least partly because of several shared effects on neural coding. We use both the FitzHugh-Nagumo (FHN) neuron model, a common model in SR research \citep{Wiesenfeld1994,Collins1995,Longtin1998,Stocks2001}, as well as the leaky integrate-and-fire (LIF) neuron model, a typical model of cortical neurons \citep{Koch1999}. We use a common aperiodic random signal $s(t)$ to drive populations of FHN and LIF neurons with varying levels of internal noise and heterogeneity, and compare the similarity of the decoded output signal with the input signal using mutual information. Stochastic offset is modeled as zero-mean Gaussian white noise with standard deviation $\sigma_\eta$, independently and identically distributed across all neurons (see \scn{methods}). Heterogeneity was added by randomly selecting the bias voltages $b_i$ from a uniform interval of varying width. We report how noise and heterogeneity both act to increase information transmission through these populations. They achieve this using three common mechanisms---linearizing the response of a population of neurons to a stimulus, desynchronizing neuronal firing, and decreasing the reaction time of a population of neurons to a stimulus---which we examine in detail.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\scnlabel{results}

We performed numerical experiments analyzing how noise and heterogeneity affect the information transmission capacity of populations of both FitzHugh-Nagumo (FHN) neurons and leaky integrate-and-fire (LIF) neurons, to test whether heterogeneity has similar effects to noise in increasing information transmission. Specifically, trials were performed using four distinct levels of random heterogeneity across the bias voltages $b_i$, the levels being none, low, medium, and high, corresponding to 0, 0.1, 0.5, and 2 times the input signal standard deviation $\sigma_s = 0.1$ (where the highest level spans the expected range of the input signal). At each level of heterogeneity, the standard deviation $\sigma_\eta$ of a stochastic offset---applied to the voltage parameter of the neuron model---was varied logarithmically from an insignificant level ($\sigma_\eta = 10^{-9}$ for FHN neurons and $\sigma_\eta = 10^{-6}$ for LIF neurons) to levels 10 times the magnitude of the input signal ($\sigma_\eta = 10^0$), in increments of $\log_{10}(\Delta\sigma_\eta) = 0.1$. For the remainder of this paper, we will use the term noise to refer specifically to this stochastic offset. At each level of heterogeneity and noise, 25 trials were performed to achieve an accurate estimate of the mean information transmission, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.

Each trial consisted of presenting a population of $N = 64$ FHN or LIF neurons with 4.5 seconds of band-limited aperiodic random signal with zero mean and standard deviation $\sigma_s = 0.1$. This signal has equal power below 5 Hz and zero power above 5 Hz, where the threshold of 5 Hz was chosen because it is significantly slower (approximately $\frac{1}{5}$) of the maximum firing rate of the FHN and LIF neurons (25--30 Hz). The population was divided into 32 ``on'' neurons and 32 ``off'' neurons, where each ``on'' (``off'') neuron received a positive (negative) version of the aperiodic input signal as input. The FHN or LIF neurons were simulated using their respective differential equations, and the output was decoded using a simple summing neuron that summed and low-pass filtered the outputs of the individual neurons to obtain the population output (outputs from ``off'' neurons were multiplied by -1 before summing). Finally, the mutual information between the aperiodic input signal and the decoded output signal was calculated. The mutual information for each level of heterogeneity and noise was computed as the average mutual information from all 25 trials at those levels. See \scn{methods} for the full details of the experimental setup.

\fig{info} shows the results of these experiments. In the homogeneous populations, a strong stochastic resonance effect is present, in which the transmitted information (\ie mutual information between the input and decoded output signals) was optimized for noise levels of $\sigma_\eta \approx 3 \times 10^{-3}$ (FHN neurons) and $\sigma_\eta \approx 1 \times 10^{-2}$ (LIF neurons). These noise levels are about 30 times (FHN neurons) or 10 times (LIF neurons) smaller than the input signal standard deviation of $\sigma_s = 0.1$. This makes them small enough not to significantly degrade the input signal, while being large enough to elicit stochastic resonance effects.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/figure1_info.eps}
  \fi
  \caption{
    \textbf{Heterogeneity increases information transfer in populations of FHN and LIF neurons.} We simulated populations of 64 FHN neurons and 64 LIF neurons, varying the level of noise ($x$-axis) and heterogeneity (individual traces), and measured the mutual information between the aperiodic input signal $s(t)$ (where $\sigma_s = 0.1$) and the output signal $r(t)$ ($y$-axis). Moderate to high levels of heterogeneity resulted in an increase in information transmission at low to moderate noise levels in both FHN and LIF neurons. The increases in FHN neurons were modest, with a strong stochastic resonance effect still exhibited at all levels of heterogeneity. The increases in LIF neurons were much more pronounced, with moderate to high levels of heterogeneity surpassing the results achieved by stochastic resonance.
  }
  \figlabel{info}
\end{figure}

For both FHN and LIF neurons, heterogeneity also increases the information transmission of the population, specifically for noise levels below the optimal stochastic resonance noise level. In FHN neurons, the benefits of adding heterogeneity to neuron biases are modest (see \scn{phase}), resulting in a significant increase in information transmission, but not achieving the same levels of information transmission as stochastic resonance (\fig{info}A). For LIF neurons, the benefits of heterogeneity in neuron biases are pronounced, with moderate to high levels of heterogeneity achieving the same or even better levels of information transmission---across a wide range of noise values---as the optimal level of noise (\fig{info}B).

In the remaining experiments, we consider three mechanisms that allow both noise and heterogeneity to improve signal transmission in populations of neurons: 1) putting neurons out of phase with one another, 2) decreasing the reaction time of a population of neurons to a stimulus, and 3) linearizing the response of a population of neurons to a stimulus. Though noise and heterogeneity affect neurons differently, they both act to achieve these three goals. None of the three factors is sufficient in and of itself to explain the effects of stochastic resonance and heterogeneity on nonlinear dynamical systems, such as the populations of FHN and LIF neurons studied. Nevertheless, by examining each factor individually, we gain a deeper understanding of how both stochastic resonance and heterogeneity operate in nonlinear dynamical systems, and ultimately show that the two concepts are fundamentally connected (see \scn{discussion}).

\subsection{Phase}
\scnlabel{phase}

Consider the following: a group of perfectly homogeneous neurons with no noise, such that their dynamics are completely deterministic. Begun with the same initial conditions, they will follow exactly the same trajectory in phase space, since they all receive the same input signal and obey the same set of differential equations. Such a population of simulated neurons will exhibit perfectly synchronized firing events, and therefore the neurons contribute redundant information: any signal that could be decoded from the firing events of this population of neurons could also be decoded from the spikes of a single neuron. These neurons are perfectly in phase: at any time, they are all in the same phase of their firing cycle, and in the same position in phase space.

Adding independent noise to each neuron desynchronizes the firing events of the population over time by causing diffusion of the neuronal phases. A small random perturbation causes a slightly advanced or delayed firing of a neuron, creating a small shift in the phase of that neuron. Specifically, a firing neuron follows a stable limit cycle attractor in phase space; perturbations tangent to the limit cycle result in diffusion along the limit cycle, while perturbations perpendicular to the limit cycle rapidly return the the stable attractor \citep{Tomita1974}. Over time, this diffusion along the limit cycle results in the population of neurons having a wide distribution of phases (\fig{phase}), and the neurons fire asynchronously.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/figure2_phase.eps}
  \fi
  \caption{
    \textbf{Noise and heterogeneity both desynchronize neurons.} To measure whether noise and heterogeneity have a desynchronizing effect on populations of neurons, we measured the standard deviation of neuron phases across the population (see \scn{methods}), where a smaller standard deviation indicates increased synchronization. Heterogeneity is seen to significantly desynchronize both FHN and LIF neurons, specifically across low to moderate levels of noise. At high levels of noise, the noise itself is sufficient to desynchronize neurons, and additional heterogeneity has no effect.
  }
  \figlabel{phase}
\end{figure}

This previous example is, of course, artificial: it requires that all neurons are \emph{exactly} the same, and start with \emph{exactly} the same initial conditions. In a more realistic scenario, one expects that small differences in neuronal parameters or small amounts of noise will cause neurons to desynchronize over a long period of time. This is not the case, because signals below the firing thresholds of a group of neurons have a synchronizing effect on the group. By stopping neuronal firing, these signals push all the neurons in the group towards their stable equilibrium points. When the signal returns to an amplitude above the firing thresholds of neurons in the group, all neurons will begin firing at approximately the same time, and will be once again in phase. Signals with higher frequency components are able to cross larger ranges of firing thresholds in quick succession, and have a stronger synchronizing effect than purely low frequency signals. Thus, more negative (positive) portions of the input signal repeatedly act to synchronize ``on'' (``off'') neurons, and groups of neurons must be able to desynchronize quickly after these synchronization events in order to remain out of phase with one another.

Similar to noise, heterogeneity acts to desynchronize neurons. Adding heterogeneity to a population significantly increases the range of distribution of neuron phases, thereby desynchronizing populations of both FHN and LIF neurons (\fig{phase}). Compared with a noiseless homogeneous population, where changes in an input signal will cross the common bias of all neurons at the same time and thereby put all neurons in phase with one another, changes in an input signal in a heterogeneous population will sequentially cross the biases of different neurons, causing some to begin firing before others. For example, an increasing input signal will cause neurons with lower biases to begin firing before those with higher biases, resulting in desynchronized firing within the population of neurons.

High frequency input signals may cross the biases of many neurons in a heterogeneous population at the same time, causing a large group of neurons to begin firing simultaneously. This effect does not have a significant synchronizing effect on LIF neurons, since the LIF tuning curves used have significantly different firing rates for different sizes of suprathreshold signal (\fig{tuning}). Specifically, a high-frequency increase in signal amplitude starts neurons firing at approximately the same time, but starts them firing at different rates, allowing them to quickly desynchronize. FHN neurons, however, all have around the same firing rate regardless of the strength of the suprathreshold signal (\fig{tuning}), and heterogeneity with higher-frequency signals has a reduced effect, since these signals cause many neurons to begin firing at once, unable to desynchronize due to their common firing rate. This discrepancy between LIF and FHN neurons results in heterogeneity having a large effect on phase in LIF neurons (\fig{phase}B), whereas the desynchronizing effect of heterogeneity in FHN neurons is much more modest (\fig{phase}A).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Response time}

Both noise and heterogeneity benefit the response time of a population of neurons, \ie how quickly the population responds to a rapidly changing signal. To measure this, we presented populations of FHN and LIF neurons with a step function and observed how well the population can follow the step function (\fig{step}). Though noise and heterogeneity both offer decreased response time, they achieve this using different mechanisms.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/figure3_step.eps}
  \fi
  \caption{
    \textbf{Noise and heterogeneity both improve the step response of neuronal populations.} We excited populations of 100 FHN neurons and 100 LIF neurons using a step function, and measured the responses of populations without noise or heterogeneity ($b_i = 0, \sigma_\eta = 0$), with noise ($b_i = 0, \sigma_\eta = 10^{-3}$), and with heterogeneity ($b_i \in [-0.01,0.01], \sigma_\eta = 0$). Each line depicts the average population response over 100 trials, and the surrounding region shows one standard deviation. FHN populations without noise or heterogeneity show a considerable delay in responding to the step function, because all neurons are close to their stable equilibrium. Both noise and heterogeneity significantly reduce this delay, and also reduce oscillations by desynchronizing neurons (\fig{phase}). LIF populations only exhibit a mild initial delay, which noise and heterogeneity both correct; in LIF neurons, the main benefit to the step response is reducing oscillations through desynchronization.
  }
  \figlabel{step}
\end{figure}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{times}
\usepackage{graphicx}
\usepackage{color}
\usepackage{multirow}

\usepackage[authoryear]{natbib}

\usepackage{rotating}
\usepackage{bbm}
\usepackage{latexsym}

%%% margins
\textheight 23.4cm
\textwidth 14.65cm
\oddsidemargin 0.375in
\evensidemargin 0.375in
\topmargin  -0.55in

\renewcommand{\baselinestretch}{2}

\interfootnotelinepenalty=10000

@@

\usepackage{nameref}

%% \newcommand{\hidefigures}{\relax}

\begin{document}
\hspace{13.9cm}

@@

{\LARGE The competing benefits of noise and heterogeneity in neural coding}

\ \\
{\bf \large Eric Hunsberger$^{\displaystyle 1}$, Matthew Scott$^{\displaystyle 2}$, Chris Eliasmith$^{\displaystyle 1}$}\\
{$^{\displaystyle 1}$Centre for Theoretical Neuroscience, University of Waterloo, Waterloo, Ontario, Canada.}\\
{$^{\displaystyle 2}$Department of Applied Mathematics, University of Waterloo, Waterloo, Ontario, Canada.}\\
%

@@

\thispagestyle{empty}
\markboth{}{NC instructions}
%
\ \vspace{-0mm}\\
%
%Abstract
\begin{center} {\bf Abstract} \end{center}

Noise and heterogeneity are both known to benefit neural coding. Stochastic resonance describes how noise, in the form of random fluctuations in a neuron's membrane voltage, can improve neural representations of an input signal. Neuronal heterogeneity refers to variation in any one of a number of neuron parameters, and is also known to increase the information content of a population. We explore the interaction between noise and heterogeneity, and find that their benefits to neural coding are not independent. Specifically, a neuronal population better represents an input signal when either noise or heterogeneity is added, but adding both does not always improve representation further. To explain this phenomenon, we propose noise and heterogeneity both operate using two shared mechanisms: 1) temporally desynchronizing the firing of neurons in the population, and 2) linearizing the response of a population to a stimulus. We first characterize the effects of noise and heterogeneity on the information content of populations of either leaky integrate-and-fire or FitzHugh-Nagumo neurons. We then examine how the mechanisms of desynchronization and linearization produce these effects, and find that they work to distribute information equally across all neurons in the population, both in terms of signal timing (desynchronization) and in terms of signal amplitude (linearization). Without noise or heterogeneity, all neurons in the population encode the same aspects of the input signal; adding noise or heterogeneity allows neuron to encode \emph{complementary} aspects of the input signal, thereby increasing information content. The simulations detailed in this paper highlight the importance of heterogeneity and noise in population coding, demonstrate their complex interactions in terms of the information content of neurons, and explain these effects in terms of underlying mechanisms.

%%%%%%%%%%%

\section{Introduction}
\label{scn:introduction}

@@

The early stochastic resonance literature dealt mainly with \emph{subthreshold} stochastic resonance \citep{Gammaitoni1998}. In a deterministic setting, systems which only respond to inputs above a certain threshold---such as neurons---cannot transmit any information about subthreshold input signals, \emph{i.e.,} input signals which are entirely below the system's threshold. In contrast, a noisy input signal (\emph{i.e,} a signal with an added stochastic offset) will occasionally cross the system's threshold, and thereby allows the system to transmit more information about the input signal. Too much noise, though, causes the input signal to become lost in random fluctuations. A moderate amount of noise---not too much, and not too little---is needed to optimize the information about the input signal contained in the system \citep{Wiesenfeld1994,Longtin1998}.

A new form of stochastic resonance---called \emph{suprathreshold} stochastic resonance (SSR)---has emerged in the past decade \citep{McDonnell2009,McDonnell2011}. In neuroscience, suprathreshold SR looks at the benefits of noise to populations of nonlinear neurons receiving a common signal of which a significant portion is above the systems' firing threshold(s). This concept was first introduced by \cite{Stocks2000}, who looked at groups of binary threshold units, and found that noise increases information transmission in such systems. Subsequent research has examined the SSR effect in populations of more sophisticated simulated neurons \citep{Stocks2001}.

The fact that neurons are heterogeneous---with variation across a wide range of parameters---has largely been ignored in the stochastic resonance literature. It is well known that neurons vary widely across the nervous system. Even neurons in the same brain area and of the same qualitative type vary across numerous parameters. Though this diversity in neural systems was documented by some of the first researchers in the field, it is only recently that neuroscientists have begun to illuminate the functional role of heterogeneity. Early work by \cite{Eliasmith2003} demonstrated that heterogeneity increases both the ability of a population to represent an input signal, and the range of functions computable by the population. More recent work showed that heterogeneity allows neurons to use combinatorial coding schemes \citep{Osborne2008} and in general increase the information encoded by a population \citep{Shamir2006,Chelaru2008,Padmanabhan2010,Ecker2011}. In the field of stochastic resonance, the effects of noise in a heterogeneous population have been examined, but generally in specific cases such as for input signals much larger than the range of heterogeneity \citep{Stocks2000}, or for very large levels of noise \citep{McDonnell2006}. Furthermore, research into stochastic resonance with heterogeneity has focused on populations of binary threshold units, which fail to capture the dynamics of actual neurons.

The purpose of this paper is to compare how noise and heterogeneity help a neuronal population to represent an input signal. Previous studies have looked at the individual benefits of noise and heterogeneity in population coding, but the interplay of these two factors has not been examined in detail. We find that noise and heterogeneity exhibit complex interactions, and though they both help to increase the information contained in a population, adding both does not always increase information content further, as compared with adding only one or the other. We argue that the similar benefits of noise and heterogeneity, as well as their complex interactions, occur because they operate on the dynamics of a neuronal population through shared mechanisms.

We explore this hypothesis in the context of two neuron models: the FitzHugh-Nagumo (FHN) neuron model, a common model in SR research \citep{Wiesenfeld1994,Collins1995,Longtin1998,Stocks2001}, and the leaky integrate-and-fire (LIF) neuron model, a typical model of cortical neurons \citep{Koch1999}. We drive the population of FHN or LIF neurons with a common aperiodic random signal $s(t)$, and vary the levels of internal noise and heterogeneity in the population. Finally, we decode an output signal from the population by summing and filtering the spikes from individual neurons, and compare the similarity of the decoded output signal with the input signal using mutual information. Stochastic offset is modeled as zero-mean Gaussian white noise with standard deviation $\sigma_\eta$, independently and identically distributed across all neurons (see \textsc{\nameref{scn:methods}}). Heterogeneity was added by randomly selecting the bias voltages $b_i$ from a uniform interval of varying width.

We report how noise and heterogeneity both act to increase information content in these populations, but in a complex, non-independent manner. They achieve this using two common mechanisms---linearizing the response of a population of neurons to a stimulus, and desynchronizing neuronal firing---that we examine in detail.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{scn:results}

We performed numerical experiments analyzing how noise and heterogeneity affect the information capacity of populations of both FitzHugh-Nagumo (FHN) neurons and leaky integrate-and-fire (LIF) neurons. Heterogeneity was added to a population by selecting the firing thresholds ${b_i}$ from a uniform distribution $\mathcal{U}(-b_r, b_r)$, where $b_r$ controls the degree of heterogeneity, with larger $b_r$ corresponding to a more heterogeneous population (on average). Noise was added to the voltage parameter of the neuron model via a stochastic offset $\eta(t)$, Gaussian distributed with zero-mean and standard deviation $\sigma_\eta$, and independent and identically distributed for each neuron. For the remainder of this paper, we will use the term noise to refer specifically to this stochastic offset. At each level of heterogeneity and noise, 25 trials were performed to achieve an accurate estimate of the mean information content, with each trial using a unique set of random biases and a unique set of stochastic offsets added to the membrane voltage.

Each trial consisted of presenting a population of $N = 64$ FHN or LIF neurons with 4.5 seconds of band-limited aperiodic random signal with zero mean and standard deviation $\sigma_s = 0.1$. This signal has equal power below 5 Hz and zero power above 5 Hz, where the threshold of 5 Hz was chosen because it is significantly slower (approximately $\frac{1}{5}$) of the maximum firing rate of the FHN and LIF neurons (25--30 Hz). The population was divided into 32 ``on'' neurons and 32 ``off'' neurons, where each ``on'' (``off'') neuron received a positive (negative) version of the aperiodic input signal as input. The FHN or LIF neurons were simulated using their respective differential equations, and the output was decoded using a simple summing neuron that summed and low-pass filtered the outputs of the individual neurons to obtain the population output (outputs from ``off'' neurons were multiplied by -1 before summing). Finally, the mutual information between the aperiodic input signal and the decoded output signal was calculated. The mutual information for each level of heterogeneity and noise was computed as the average mutual information from all 25 trials at those levels. See \textsc{\nameref{scn:methods}} for the full details of the experimental setup.

Figure~\ref{fig:infohetero} summarizes the effect of neuronal heterogeneity on the information content of a population of LIF neurons. At low noise levels, we observe a resonance effect in the heterogeneity parameter---there is an optimal level of heterogeneity that maximizes the information content of the population. Furthermore, this optimal level of heterogeneity is around one to two times the root mean square amplitude of the input signal. This indicates that optimal heterogeneity corresponds to having neuron firing thresholds that span the full range of input signal values. Too little heterogeneity means that some input signal components cannot be uniquely encoded by the population, since these values will be either well below or well above the firing thresholds of all neurons in the population, and all neurons will fire approximately constantly for these values. Too much heterogeneity results in some neurons that have firing thresholds well below or well above all signal values, causing them to either fire tonically independent of the input signal value, or to fire for no part of the input signal, respectively. This type of resonance in heterogeneity was recently observed by \cite{Mejias2012}, both theoretically and in numerical experiments.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics{../figures/bw/infohetero.eps}
  \fi
  \caption{
    \textbf{Heterogeneity increases information content of a population of LIF neurons, and exhibits a resonance effect.}
    Adding heterogeneity to a population of LIF neurons increases their ability to represent an input signal, but only for lower levels of noise. At higher levels of noise, the noise alone is able to fully desynchronize neuronal firing and linearize neuron responses, so adding heterogeneity offers no additional benefit.
  }
  \label{fig:infohetero}
\end{figure}

At higher levels of noise, heterogeneity offers no benefit to LIF neuron populations (Figure~\ref{fig:infohetero}). To examine the interactions between noise and heterogeneity further, we conducted the same experiment across a wide range of noise and heterogeneity values. Figure~\ref{fig:infocontour} illustrates the effects of these interactions on the information content of a population of LIF or FHN neurons. In LIF neurons, the population is best able to represent the input signal with a moderate amount of heterogeneity ($b_r \approx 1.5 \times 10^{-1}$) and no noise ($\sigma_\eta = 1.0 \times 10^{-4}$). Given this optimal level of heterogeneity, adding noise only serves to decrease the information content of the system. But if no heterogeneity is present in the system, then adding the right amount of noise ($\sigma_\eta \approx 1.0 \times 10^{-2}$) optimizes information content (Figure~\ref{fig:infonoise}). This shows that the effects of noise and heterogeneity are not independent.

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infocontour.eps}
  \fi
  \caption{
    \textbf{Noise and heterogeneity show nonlinear interactions in their benefits to information representation.} We measured the mutual information (contours every 0.15 bits) across many levels of noise and heterogeneity. In LIF neurons, both noise and heterogeneity show independent resonance effects: given that only one of noise or heterogeneity is present, there is a non-zero and finite level of noise or heterogeneity that optimizes information content. Setting both noise and heterogeneity to their optimal levels fails to further improve the representational ability of the population. Both parameters use the same underlying mechanisms of desynchronization and linearization, so once the population has been desynchronized and linearized by one parameter, adding the other cannot further improve information content. Heterogeneity is unable to fully desynchronize FHN neurons (see \textsc{\nameref{scn:desync}}), so information content in these neurons is maximized for a particular level of noise.
  }
  \label{fig:infocontour}
\end{figure}

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/infonoise.eps}
  \fi
  \caption{
    \textbf{Noise increases information content of homogeneous populations of FHN and LIF neurons.} In both LIF (left) and FHN (right) populations, noise exhibits a stochastic resonance effect, where adding noise to the homogeneous population (black line) improves the population's information content. This stochastic resonance effect disappears in heterogeneous populations of LIF neurons, but still occurs in heterogeneous populations of FHN neurons.
  }
  \label{fig:infonoise}
\end{figure}

We propose that noise and heterogeneity both improve the information representation ability of a population of neurons by 1) desynchronizing the firing of neurons in the population, and 2) linearizing the response of the population to an input signal. These shared mechanisms result in the nonlinear interaction between noise and heterogeneity in terms of information content: once either noise or heterogeneity has adequately desynchronized a population, the addition of the other cannot further desynchronize or linearize, and therefore provides no additional increase in information representation. In the following two sections, we present additional experimental results to investigate and quantify the mechanisms of desynchronization and linearization.

We also note that there is a difference between FHN and LIF neurons in terms of noise and heterogeneity, and their effects on information content. Heterogeneity is much less effective at increasing information content in FHN neurons than LIF neurons (Figures~\ref{fig:infocontour}~and~\ref{fig:infonoise}). This is explained by the inability of heterogeneity to fully desynchronize firing in FHN neurons (see \textsc{\nameref{scn:desync}}), which is in turn due to the nature of the FHN tuning curves. We revisit this result in the \textsc{\nameref{scn:discussion}}.

\subsection{Desynchronization}
\label{scn:desync}

Consider the following: a group of perfectly homogeneous neurons with no noise, such that their dynamics are completely deterministic. Begun with the same initial conditions, they will follow exactly the same trajectory in phase space because they all receive the same input signal and obey the same set of differential equations. Such a population of simulated neurons will exhibit perfectly synchronized firing events, and therefore the neurons contribute redundant information. Any signal that could be decoded from the firing events of this population of neurons could also be decoded from the spikes of a single neuron. These neurons are perfectly synchronized; at any time, they are all in the same phase of their firing cycle, and in the same position in phase space.

Adding independent noise to each neuron desynchronizes the firing events of the population by causing diffusion of the neuronal phases \citep{Stocks2001a}. A small random perturbation causes a slightly advanced or delayed firing of a neuron, creating a small shift in the phase of that neuron. Specifically, a firing neuron follows a stable limit cycle attractor in phase space; perturbations tangent to the limit cycle result in diffusion along the limit cycle, while perturbations perpendicular to the limit cycle rapidly return the the stable attractor \citep{Tomita1974}. Over time, this diffusion along the limit cycle results in the population of neurons having a wide distribution of phases (Figure~\ref{fig:phase}), and the neurons fire asynchronously (Figure~\ref{fig:syncraster}).

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/phase.eps}
  \fi
  \caption{
    \textbf{Noise and heterogeneity both desynchronize neurons.} To measure whether noise and heterogeneity have a desynchronizing effect on populations of neurons, we measured the standard deviation of neuron phases across the population (see \textsc{\nameref{scn:methods}}), where a smaller standard deviation indicates increased synchronization. Heterogeneity is seen to significantly desynchronize both FHN and LIF neurons, specifically across low to moderate levels of noise. At high levels of noise, the noise itself is sufficient to desynchronize neurons, and additional heterogeneity has no effect.
  }
  \label{fig:phase}
\end{figure}

\begin{figure}
  \ifx\hidefigures\undefined
    \centering
    \includegraphics[width=\textwidth]{../figures/bw/syncraster.eps}
  \fi
  \caption{
    \textbf{The effects of noise and heterogeneity on LIF neuron firing.}
    Spike rasters illustrate the effects of noise and heterogeneity parameters on neuronal firing in a population of 64 LIF neurons. With low noise and heterogeneity, neuronal firing is highly synchronized. Adding a moderate amount of noise or heterogeneity desynchronizes this firing. Adding a moderate level of both parameters cannot provide further desynchronization, since firing is already quite desynchronized. High levels of noise cause rapid firing in almost all neurons, degrading the ability of the population to represent the input signal. High levels of heterogeneity cause the signal to be entirely above or below the firing thresholds of many neurons, again degrading representational ability. Noise: $\sigma_\eta = $$10^{-3}$ (low), $10^{-2}$ (moderate), $10^{-1}$ (high). Heterogeneity: $b_r = $$10^{-2}$ (low), $10^{-1}$ (moderate), $10^{0}$ (high).
  }
  \label{fig:syncraster}
\end{figure}

This previous example is, of course, artificial. It requires that all neurons are \emph{exactly} the same, and start with \emph{exactly} the same initial conditions. In a more realistic scenario, one expects that small differences in neuronal parameters or small amounts of noise will cause neurons to desynchronize over a long period of time. This is not the case, because signals below the firing thresholds of a group of neurons have a synchronizing effect on the group. By stopping neuronal firing, these signals push all the neurons in the group towards their stable equilibrium points. When the signal returns to an amplitude above the firing thresholds of neurons in the group, all neurons will begin firing at approximately the same time, and will be once again in phase. Signals with higher frequency components are able to cross larger ranges of firing thresholds in quick succession, and have a stronger synchronizing effect than purely low frequency signals. Thus, more negative (positive) portions of the input signal repeatedly act to synchronize ``on'' (``off'') neurons, and groups of neurons must be able to desynchronize quickly after these synchronization events in order to remain out of phase with one another.
